---
title: "Covid19 vaccine study Final2021"
author: "Taylor Stohl"
date: "Summer 2021"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---
 
```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
#FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```




# Introduction

A new preventative vaccine (covid-19) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of sub populations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.


# Coding details 

First, the variables from FinalData needed to be changed from character variables into factor variables. the appropriate dplyr code was used to do this. The following code was used to achieve this: 

FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))

Next, specific subgroups had to be separated from the broader categories in order to properly studied. Gal and guy were separated into subgroups, as well as gay sexual orientation and people that answered  Druguser to participating in drug use. All of these categories are isolated to see weather or not these particular populations should  or should not take the  new covid-19 vaccine.




# Guys 

### Barplot



```{r}
dd2 <- guy %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```




This plot represents the population of guys(males) people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 


The plot above shows that a shocking 58% of the people that took the vaccine did in fact get covid-19, while only approximately 42% of the people taking the placebo contracted covid-19.  A substantially higher percentage of people who took tha vaccine contracted covid-19 versus those who were only given the placebo. The vaccine appears to be ineffective in protection against covid-19. The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of guys(men) that got covid-19 after taking the placebo and taking the vaccine. 
Among the guys, 58% got covid-19 after taking the vaccine, while the approximately 42% of guys that were given the placebo contracted covid-19. The vaccine is clearly not effective in preventing covid-19 in guys(males) and it is not recommended. Those who took the placebo were much less likely to contact covid-19
The odds of getting covid-19 while taking the vaccine is 3:2. The men that took the drug were one and half times more likely to get covid-19 than those taking the placebo.


### Inferential 



```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```


The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 6e-04. This is an awfully small number. seeing this number means that these results are not likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.



The null hypothesis predicted that the deaths related to covid-19 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covi-19 deaths and sex, is the favorable option.

Like the odd ratio above illustrates, guys(men) who recieved tha vaccine were 1.5 times more likely to get covid-19 than those who received the placebo. 


```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)
print(effectivenessguy*100)
```

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`




# Gals 


### Barplot

```{r}
dd2 <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This plot represents the population of gals(females) people and the relationship to the contraction of covid-19 by either talking the vaccine or the placebo. 

In this plot, contrary to the previous, it appears that more people who were only given the placebo contracted covid-19. Only approximately 33% of the individuals who were given the vaccine were reported to have contracted covid-19. That means around 67% of the people that got covid-19 were the individuals that only received the placebo.The odds of getting covid-19 while taking the vaccine is 1:2. The gals that took the vaccine were half as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 



The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of gals(women) that got covid-19 after taking the placebo and taking the vaccine. 
The gals(women) that took the vaccine ans contracted covid-19 was 33%, just as approximated in the plots prior. The gals that took the placebo and got covid-19 was 66%. The vaccine is in fact effective in protecting gals against the contraction of covid-19. 
The odds of getting covid-19 while taking the vaccine is 1:2 for gals. The gals that took the vaccine were half as likely to get covid-19 than those taking the placebo.



### Inferential 




```{r}
chisq.test(table1)
chisqtestGC(table1)
galsfish <- fisher.test(table1)
galsfish
```



The information to play close attention to among all of these results is the output of the Fisher's Exact test. The p-value displayed above is extremely small at 0.0000000000000002. This is an extremely tiny number. seeing this number means that these results are not likely to occur by chance and so the null hypothesis can be rejected in favor of the alternative.


The Fishers's Exact Test illustrates that the odds ratio is 0.49.
gals(Women) who got the drug were about 0.49 times more likely to get covid-19 than women that only took the placebo.


```{r}
effect <- galsfish$p.value/(galsfish$p.value +1)
effectivenessgal = guysfish$estimate/(galsfish$estimate +1)
print(effectivenessgal*100)
```

```{r}
low=(galsfish$conf.int[1]/(galsfish$conf.int[1] +1))
hi=(galsfish$conf.int[2]/(galsfish$conf.int[2] +1))
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`






# Gay Sexual Orientation

### Barplot

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```



This plot represents the population of gay people and the relationship to the contraction of covid-19 by taking either the vaccine or the placebo.

In this barplot, it appears that a sniffling 71% of the individuals that took the vaccine got covid-19 as well as the 29% who received the placebo. The odds of getting covid-19 while taking the vaccine is 2.5:1. The individuals who identified as gay and took the vaccine were three times as likely to get covid-19 than those taking the placebo.


There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 


The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

These numerical results show specific percentages of gay individuals that got covid-19 after taking the placebo and taking the vaccine. 
The gay participants that took the vaccine and contracted covid-19 was at an approximately shocking 71%! Those participants that only received the placebo only contacted covid-19 at 29%. The vaccine id evidently completely ineffective in protecting the gay community against covid-19 and taking it is not advised. 
The odds of getting covid-19 while taking the vaccine is 2.5:1. The individuals who identified as gay and took the vaccine were three times as likely to get covid-19 than those taking the placebo.





### Inferential 




```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```


The p-value is 6e-06. This in an incredibly small number and with that being said, the conclusion is that these results are unlikely to occur by chance.

The Fishers's Exact Test illustrates that the odds ratio is 2.5.
Gay individuals who got the vaccine were about 2.5 times more likely to get Covid-19 than gays that took the placebo.

The null hypothesis is rejected in favor the the alternative. The drug is completely ineffective in preventing the chances of getting covid-19 in individuals who identify as gay.





```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgays = gaysfish$estimate/(gaysfish$estimate +1)
print(effectivenessgays*100)
```

```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))
options(digits=2)
```





## Effectiveness Confidence Interval `r low`  to `r hi`





# People that are Drugusers

### Barplot


```{r}
dd2 <-  Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This plot represents the population of people who answered  Druguser to drug use and the relationship to the contraction of covid-19 while taking either the vaccine or the placebo.

This barplot shows that arounf 34% of the people who answered yes to drug use and recieved the vaccine contracted covid-19. This means that around 66% of the drugusers that only took the placebo eded up getting covid-19.The odds of getting covid-19 while taking the vaccine is 1:2. The drugusers that took the vaccine were half as likely to get covid-19 than those taking the placebo.




There are two plots. In first plot,the x-axis represents infected, whether the participants were infected with covid-19 or they were not. The y-axis represents the count, or number of people that got covid-19 or not.
In the second plot, the one above, The The left two bars of the graph are all the people that got covid-19. The x-axis represents whether a person got covid-19 or not. The y-axis, however, instead of representing the number of people who got covid-19, now represents the percentage of people that got covid-19 after taking the the placebo or taking the Vaccine. 




The codes that  create this plot start with the code block beginning with dd2. dd2 can be seen in the environment pane and if inspected contains 4 variables.These variables include infected, treatment, count, and prcent(stands for percent). The second group of codes is called the basicC. BasicC can also be seen in the environment pane and appears to contain a list of 9 variables. This information creates the first plot. The basicCC is the last bunch of codes and is also in the files pane. It also contains 9 variables and helps to create the percentage plot, the second plot.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```
 

These numerical results show specific percentages of drugusers that got covid-19 after taking the placebo and taking the vaccine.
The percentage of drug users that took the vaccine and contacted covid-19 is 33% Those that were only given the placebo and got covid-19 was almost 67%. It is clear that the vaccine to prevent covid-19 is effective in protecting drug users against covid-19 and  so getting this vaccine is recommended. The odds of getting covid-19 while taking the vaccine is 1:2. The drugusers that took the vaccine were half as likely to get covid-19 than those taking the placebo.


### Inferential 


```{r}
chisq.test(table1)
chisqtestGC(table1)
drugfish <- fisher.test(table1)
drugfish
```


The p-value is 0.002. This in a small number and with that being said, the conclusion is that these results are unlikely to occur by chance.

The Fishers's Exact Test illustrates that the odds ratio is 0.55.
drug users who got the vaccine were about half as likely to get Covid-19 than drug users that took the placebo.

The null hypothesis is rejected in favor the the alternative. The drug is  effective in preventing the chances of getting covid-19 in drugusers.




```{r}
effect <- drugfish$p.value/(drugfish$p.value +1)
effectivenessdrug = drugfish$estimate/(drugfish$estimate +1)
print(effectivenessdrug*100)
```

```{r}
low=(drugfish$conf.int[1]/(drugfish$conf.int[1] +1))
hi=(drugfish$conf.int[2]/(drugfish$conf.int[2] +1))
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`


# Conclusion and recommendations 

The covid-19 vaccine was not effective in protecting men against the contraction of the disease. The covid-19 effectiveness was `r effectivenessguy`

The covid-19 vaccine was successful in protecting women against covid-19. the odds ratio was 1:2.
The covid-19 vaccine was not effective in treating people who identified as gay. The odds of getting covid-19 while taking the vaccine is 3:1. The individuals who identified as gay and took the vaccine were three times as likely to get covid-19 than those taking the placebo.
The individuals that were drugusers were effectively protected by the vaccine. .The odds of getting covid-19 while taking the vaccine is 1:2. The drugusers that took the vaccine were half as likely to get covid-19 than those taking the placebo.

If you are a male or identify as gay it is not recommended to recieve the vaccine. However, if you are a female or a druguser, the vaccine was reported to have been effective.



